Comparison of several sequential and parallel derivative-free global optimization algorithms

### 2018-12-03 2019/PAVT/pavt_paper.tex
PAVT: update pdf

In the present paper, several sequential and parallel global optimization algorithms were considered.
A comparison of efficiencies of these ones has been done on a set of test problems. The results allow
making the following conclusions:
\begin{itemize}
  \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-access
implementations of which are available;
  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
speed and in reliability. It is manifested especially strongly on more complex multiextremal
problems;
  \item the parallel version of the Globalizer system demonstrates good speedup
when running on several nodes, on each of which a single objective function value per iterations is
computed. When solving the problems with fast computable objective functions and using multiple
threads on the nodes, the speedup for the Globalizer system degrade with increasing
number of nodes;
  \item the MIDACO system is the most suitable for simple global optimization problems with the
fast-computable objective functions. In this case, MIDACO is reliable enough and provides a linear
speedup with increasing number of nodes and threads executed on these ones in parallel.
\end{itemize}

>>>>

In the present paper, several sequential and parallel global optimization algorithms were considered.
A comparison of efficiencies of these ones has been done on a set of test problems. The results allow
making the following conclusions:
\begin{itemize}
  \item AGS\(l\) method implemented in the Globalizer system has demonstrated the convergence
speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
implementations of which are available;
  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
speed and in reliability. It is manifested especially strongly on more complex multiextremal
problems;
  \item the parallel version of the Globalizer system demonstrates good speedup
when running on several nodes, on each of which a single objective function value per iterations is
computed. When solving the problems with fast computable objective functions and using multiple
threads on the nodes, the speedup for the Globalizer system degrade with increasing
number of nodes;
  \item the MIDACO system is the most suitable for simple global optimization problems with the
fast-computable objective functions. In this case, MIDACO is reliable enough and provides a linear
speedup with increasing number of nodes and threads executed on these ones in parallel.
\end{itemize}
